---
title: "MSiA 400 Lab 1"
author: "Ruixiang Fan"
date: "10/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
# a.)

```{r}

# Read markov100.txt and create Transition Matrix
markov <- read.table("markov100.txt")
P <- as.matrix(markov)

```

```{r}

# Create the initial vector
a1 <- c(1)
i <- 1
b1 <- 0
while (i < 100) {
  a1 <- c(a1, b1)
  i = i + 1
}

```

```{r}

# Calculate the prob distribution after 10 steps
library(expm)
prob10 <- a1 %*% (P %^% 10)

# Index the fifth element of the first row representing probability of being in state 5 after 10 steps
prob10[1,5]

```

#### The probability of being in State 5 after 10 transitions is 0.045091 given we are currently in State 1.

# b.)

```{r}

# Create the initial vector
a2 <- c(1/3, 1/3, 1/3)
i <- 1
b2 <- 0
while (i < 98) {
  a2 <- c(a2, b2)
  i = i + 1
}

```

```{r}

# Calculate the prob distribution after 10 steps
library(expm)
prob10 <- a1 %*% (P %^% 10)

# Index the 10th element of the first row representing probability of being in state 10 after 10 steps
prob10[1,10]

```

#### The probability of being in State 5 after 10 transitions is 0.08126983 given we are currently at one of States 1, 2, and 3 with equal probabilities.


# c.)

```{r}

# Create the vector to replace the last row of t(P) - I matrix
i <- 1
a3 <- c(1)
b3 <- 1
while (i < 100) {
  a3 <- c(a3, b3)
  i = i + 1
}

# Create the rhs vector
j <- 1
a_rhs <- c(0)
b_rhs <- 0
while (j < 99) {
  a_rhs <- c(a_rhs, b_rhs)
  j = j + 1
}
a_rhs <- c(a_rhs, 1)

```

```{r}

# Calculate Pi Vector
Q <- t(P) - diag(100)
Q[100,] <- a3
rhs <- a_rhs
Pi <- solve(Q) %*% rhs

# Index the first element of the first column representing the steady state probability of being in State 1
Pi[1,1]

```

#### The steady state probability of being in State 1 is 0.01256589.

# d.)

```{r}

# Create the vector of 1's
a4 <- c(1)
b4 <- 1
i <- 1
while (i < 99) {
  a4 <- c(a4, b4)
  i = i + 1
}

```


```{r}

# Calculate the vector containing the mean first passage time to State 100 from State 1,2,...,99
B <- P[-100,-100]
Q4 <- diag(99) - B
e <- a4
m <- solve(Q4) %*% e

# Index the first element representing the mean first passage time from State 1 to State 100
m[1,1]

```

#### The mean first passage time from State 1 to State 100 is 254.9395.

# Problem 2
# a.)

```{r}

# Read markov100.txt and create Matrix
web <- as.matrix(read.table("webtraffic.txt", header = TRUE))

# Calculate the sum of visits for each path
webtotal <- colSums(web)

# Construct and Display matrx Traffic
Traffic <- matrix(webtotal, ncol = 9, byrow = TRUE)
Traffic

```

# b.)

```{r}

# Divide total traffic by 1000 (the number of visitors) to get transition probability
Traffic[9,1] <- 1000
P <- Traffic/1000
P

```

# c.)

```{r}

# Calculate and display Pi Vector
Q <- t(P)-diag(9)
Q[9,] <- c(1,1,1,1,1,1,1,1,1)
rhs <- c(0,0,0,0,0,0,0,0,1)
Pi <- solve(Q) %*% rhs
Pi

```

# d.)

```{r}

StayTime <- c(0.1,2,3,5,5,3,3,2,0)
Average_Time <- StayTime %*% Pi
Average_Time

```

#### The average time a visitor spend on the website until she leaves is about 1.86 minutes.

# e.)

```{r}

# Modify the matrix Traffic to indicate total traffic changes after adding the links
Traffic[2,6] <- Traffic[2,3]*0.3
Traffic[2,3] <- Traffic[2,3]*0.7
Traffic[2,7] <- Traffic[2,4]*0.2
Traffic[2,4] <- Traffic[2,4]*0.8
P <- Traffic/1000

# Calculate the Pi2 vector using the modified matrix Traffic
Q <- t(P)-diag(9)
Q[9,] <- c(1,1,1,1,1,1,1,1,1)
rhs <- c(0,0,0,0,0,0,0,0,1)
Pi2 <- solve(Q) %*% rhs
Pi2

```

#### Pi[3] = 0.1594 drops to Pi2[3] = 0.1511, Pi[4] = 0.1300 drops to Pi2[4] = 0.1108 after adding the links. Therefore, the link helped balancing the traffic. 
